Regression

### 2015-08-25 rand_forest.rmd
random forest start

library(devtools)
library('ggbiplot')
library('fpc')
library('vegan')
library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('corrgram')

@@

```{r echo=FALSE}
some <- all7d %>% select(return.pdol, mlsto, inv.mlsto, timespan,no.users,Discipline,pc.pro,
                         client.totinv,code.director,Num.disc,Business,Biz.type,
                         code.contact,code.client,Project.Value,JD.Second,Post.Code,code.ProjEng,Billing.Type,majority.pos,
                         pc.majpos)

#heatmap of any similar NA value ranges

experiment<- sapply(some, function(x) is.na(x))
NA.plot<- melt(experiment)
NA.plot$mlsto<- rep(all7d$mlsto, ncol(experiment))
#reorder levels for ggplot by total number of NA TRUE
# table(NA.plot$Var2, NA.plot$value)[,'TRUE']
NA.plot <- within(NA.plot, 
                         Var2 <- factor(Var2, 
                                               levels=names(sort(table(NA.plot$Var2, NA.plot$value)[,'TRUE'], 
                                                                 decreasing=FALSE))))

ggplot(NA.plot, aes(x=Var2, y=mlsto)) +
        geom_tile(aes(fill = value)) + 
        theme(axis.text.x=element_text(angle=45,hjust=1))

Will address each of the variables with many nA separately. 8 of these variables totally. run a separate lm for each of these

Need to transform some numeric variables:

return.pdol - cube root
inv.mlsto - log
timespan - cube root
pc.pro - leave alone
client.totinv - log

Project.Value - log
pc.majpos - log

#write a function for testing transformations

trans<- function(var= 'timespan', df= all7d, transform = 'cbrt'){
        #initial q plot
        init<- qqPlot(df[,var], 
                       main = 'untouched variable')
        print(init)

        #make trial dataframe
        trial = sapply(df[,var], function(x){
                if(transform == 'log'){
                        log(x)
                } else

                if(transform == 'neglog')
                        {
                        sign(x)*log(abs(x) + 1)
                } else

                if(transform == 'sqrt')
                {
                        sqrt(x)
                } else
                        if(transform == 'cbrt')
                {
                        sign(x) * abs(x)^(1/3)
                } else
                { x^2 }

                }) %>% 
                as.data.frame

        colnames(trial)[names(trial) %in% '.']<-'trans.var'

        #qplot for transformation
        qqtrans<- qqPlot(trial$trans.var, 
                         main = 'transformed')
        print(qqtrans)

        ##move on to density plots

        #original variable
        dens<- ggplot(df, aes_string(x=var)) + geom_histogram(aes(y=..density..))
        print(dens)

        #transformed variable
        dens.trans<- ggplot(trial, aes_string(x='trans.var')) + geom_histogram(aes(y=..density..))
        print(dens.trans)
}

trans(var= 'timespan', df= all7d, transform = 'cbrt')
trans(var= 'inv.mlsto', df= all7d, transform = 'log')
trans(var= 'return.pdol', df= all7d, transform = 'cbrt')
trans(var= 'pc.majpos', df= all7d, transform = 'log')
trans(var= 'client.totinv', df= all7d, transform = 'log')
trans(var= 'Project.Value', df= all7d, transform = 'log')

Have now decided on transformations for numeric variables.

Create all7d ready for linear regression analysis

math.cbrt <- function(x) {
        sign(x) * abs(x)^(1/3)
}

all7d$timespan.cbrt<- math.cbrt(all7d$timespan)
all7d$inv.mlsto.log <- log(all7d$inv.mlsto)
all7d$client.totinv.log <- log(all7d$client.totinv)
all7d$Project.Value.log <- log(all7d$Project.Value)
all7d$pc.majpos.log <- log(all7d$pc.majpos)

Root mean square error as initial check. Write code which transforms response variable back to original format by cubing, then finds RMSE.

@@

Now fit full variables to linear regression

@@

full = lm(return.pdol ~ 
                  inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + client.totinv.log + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng,
          data = all7d)
summary(full)

#have a look at the residuals

res<- resid(full)
plot(all7d$inv.mlsto.log, res)
abline(0,0)

plot(density(res))
qqPlot(res)

#compare some of the answers
ans<- data.frame(answer= all7d$return.pdol, full.lm = predict(full, all7d, interval= 'predict'))

RMSE(ans$answer, ans$full.lm.fit)

@@

RMSE is 0.4768 which isn't great in terms of predicting return per dollar.
R^2 is 0.1905

Let's try deleting some full variables

```{r, echo=FALSE}
full.2<- update(full, .~. -client.totinv.log)
summary(full.2)
full.3<- update(full.2, .~. -code.director)
summary(full.3)
full.4<- update(full.3, .~. -Num.disc)
summary(full.4)
full.5<- update(full.4, .~. -code.ProjEng)
summary(full.5)
full.6<- update(full.5, .~. -Biz.type + Discipline)
summary(full.6)

#RMSE check

ans<- data.frame(answer= all7d$return.pdol, full.lm = predict(full.6, all7d, interval= 'predict'))

RMSE(ans$answer, ans$full.lm.fit)

Feels better. That leaves us with

* inv.mlsto.log
* Discipline
* timespan.cbrt
* no.users
* pc.pro
* Discipline
* Business

RMSE = 0.48266

Going back to full variables, lets try pairwise multiplying..

```{r, echo=FALSE}
#try diff pairs

pairs.lm<- lm(return.pdol ~ inv.mlsto.log + Discipline + timespan.cbrt + no.users + pc.pro + Business + 
                      inv.mlsto.log*pc.pro + inv.mlsto.log*Business,
              data=all7d)

summary(pairs.lm)
pairs<- predict(pairs.lm, all7d)
ans<- data.frame(answer= all7d$return.pdol, pairs.lm = pairs)

RMSE(ans$answer, ans$pairs.lm)

#down to 0.4759

@@

addone<- function(var= 'JD.Second', df= all7d){
        reduced = df[!(is.na(df[,var])),]
        formula = paste("return.pdol ~ inv.mlsto.log + timespan.cbrt + no.users + Discipline + pc.pro + inv.mlsto.log*pc.pro + inv.mlsto.log*Business", var, sep=" + ")
        add.one = lm(as.formula(formula), data = reduced)

>>>>

library(devtools)
library('knitr')
library('ggplot2')
library("plyr")
library("dplyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('corrgram')
library('randomForest')
library('caret')

@@

Will address each of the variables with many NA separately. 8 of these variables totally. run a separate lm for each of these

No need to transform numeric variables - not required for trees

Root mean square error as initial check.

@@

Now fit full variables to random forest, tune mtry and number of trees

Try rpart first

@@

train_ind <- sample(seq_len(nrow(all7d)), size = nrow(all7d)*2/3)

train <- all7d[train_ind, ]
test <- all7d[-train_ind, ]

#create tree with training data only
tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.001))
summary(tree.all7d)

#have a look at cp values, automatic cross validation in rpart
printcp(tree.all7d)
plotcp(tree.all7d)
#min cp = 0.013 will give simplest tree within 1 std of x-val rel error
tree.all7d<- rpart(return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=train, control = rpart.control(cp= 0.011))

split.fun <- function(x, labs, digits, varlen, faclen){
  # replace commas with spaces (needed for strwrap)
  labs <- gsub(",", " ", labs)
  for(i in 1:length(labs)) {
    # split labs[i] into multiple lines
    labs[i] <- paste(strwrap(labs[i], width=25), collapse="\n")
    }
  labs
  }

prp(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
#make prediction with current tree
tree.pred<-predict(tree.all7d, test)
#calculate RMSE for test data
RMSE(test$return.pdol, tree.pred)
# 51.4 cents, already close to regression..

#try to make a rainbow tree
heat.tree <- function(tree, low.is.green=FALSE, ...) { # dots args passed to prp
  y <- tree$frame$yval
  if(low.is.green)
    y <- -y
  max <- max(y)
  min <- min(y)
  cols <- rainbow(99, end=.36)[
    ifelse(y > y[1], (y-y[1]) * (99-50) / (max-y[1]) + 50,
           (y-min) * (50-1) / (y[1]-min) + 1)]
  prp(tree, branch.col=cols, box.col=cols, ...)
  }

heat.tree(tree.all7d, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)

@@

RMSE is 0.514 which is worse than regression in predicting return per dollar.

Decision trees not affected by too many variables - leave all in

Decision trees cannot handle pairwise interactions

Let's try 10 fold cross validation and have a look at the error.
Use caret package

```{r, echo=FALSE}
#10 fold cross validation using caret package
#specify 10 fold
tc <- trainControl("cv",10)
#keep cp static at 0.011
rpart.grid <- expand.grid(.cp=0.011)
train.rpart <- train( return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng, data=all7d, method="rpart",trControl=tc,
                  tuneGrid=rpart.grid)

#RMSE = 0.50386
# Rsquared = 0.0844, quite bad

@@

addone<- function(var= 'JD.Second', df= all7d){
        reduced = df[!(is.na(df[,var])),]
        formula = paste("return.pdol ~ inv.mlsto + timespan + no.users + Discipline + pc.pro + 
                     client.totinv + code.director + 
                  Num.disc + Business + Biz.type + code.ProjEng", var, sep=" + ")
        add.one = rpart(as.formula(formula), data = reduced)

        #plot decision tree
        print(
          heat.tree(add.one, extra = 1, split.fun= split.fun, tweak=1, faclen =5, branch.type= 5)
          )

        #perform 10 fold xval using caret package
        tc = trainControl("cv",10)
        #keep cp static at 0.011
        rpart.grid = expand.grid(.cp=0.011)
        train.rpart = train(as.formula(formula), data=df, method="rpart",trControl=tc,
                  tuneGrid=rpart.grid)

        print(
        train.rpart$results
        )
